---
title: "Mental Health Services Data Set Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
    theme: cerulean
runtime: shiny
resource_files:
- Clean data for dashboard/MHSDS_ED_pooled.csv
- Clean data for dashboard/MHSDS_main_pooled.csv
- Clean data for dashboard/NHS workforce statistics - data.xlsx

---

```{r setup, include=FALSE}
#Load packages
library("tidyverse")
library("lubridate")
library("here")
library("data.table")
library("plotly")
library("hrbrthemes")
library("flexdashboard")
library("readxl")
library("DescTools")

#File locations
source(here::here("0. File locations.R"))

#Load main performance data
MHSDS_main_pooled_dashboard <- fread("Clean data for dashboard/MHSDS_main_pooled.csv",header=TRUE, sep=",", check.names=T)

  #Months
months <- MHSDS_main_pooled_dashboard %>% arrange(desc(end_date)) %>% pull(month_year) %>% unique(.)
latest_month <- MHSDS_main_pooled_dashboard %>%
  slice_max(.,end_date,n=1,with_ties=FALSE) %>% 
  pull(month_year)

#Load ED data
MHSDS_ED_pooled_dashboard <- fread("Clean data for dashboard/MHSDS_ED_pooled.csv",header=TRUE, sep=",", check.names=T)

#Load workforce data
NHS_workforce_doctors <- read_excel("Clean data for dashboard/NHS workforce statistics - data.xlsx")
NHS_workforce_other <- read_excel("Clean data for dashboard/NHS workforce statistics - other - data.xlsx")

  #Locations
locations <- MHSDS_main_pooled_dashboard %>%
  filter(.,month_year==latest_month) %>%
  filter(.,str_detect(PRIMARY_LEVEL_DESCRIPTION,"CCG")) %>%
  pull(PRIMARY_LEVEL_DESCRIPTION) %>%
  unique(.) %>%
  c("England",.)

#Reactive main performance table (place)
MHSDS_main_reac <- reactive({
  tm <- MHSDS_main_pooled_dashboard %>%
    filter(.,PRIMARY_LEVEL_DESCRIPTION==input$location) %>%
    mutate(.,start_date=lubridate::ymd(start_date),
           end_date=lubridate::ymd(end_date))
  return(tm)
})

#Reactive main performance table (place and time)
MHSDS_main_reac_both <- reactive({
  tm <- MHSDS_main_pooled_dashboard %>%
    filter(.,PRIMARY_LEVEL_DESCRIPTION==input$location,
               month_year==input$date) %>%
    mutate(.,start_date=lubridate::ymd(start_date),
           end_date=lubridate::ymd(end_date))
  return(tm)
})

#Reactive ED table (place)
MHSDS_ED_reac <- reactive({
  tm <- MHSDS_ED_pooled_dashboard %>%
    filter(.,PRIMARY_LEVEL_DESCRIPTION==input$location) %>%
    mutate(.,start_date=lubridate::dmy(REPORTING_PERIOD_START),
         end_date=lubridate::dmy(REPORTING_PERIOD_END))
  return(tm)
})

#Graph parameters
m <- list(
  l = 20,
  r = 20,
  b = 20,
  t = 20,
  pad = 2
)
```

Sidebar {.sidebar}
=======================================================================

### MHSDS parameters

This dashboard displays key statistics produced by NHS Digital from their [MHSDS](https://digital.nhs.uk/data-and-information/data-collections-and-data-sets/data-sets/mental-health-services-data-set) data series. This version is for teaching purposes only.

Application author: [Sebastien Peytrignet](https://github.com/sg-peytrignet), [The Health Foundation](www.health.org.uk)


```{r}
#Location
selectInput('location', label = 'Select a location',
            choices = locations, selected = "England")

#Date
selectInput('date', label = 'Select a month',
            choices = months, selected = latest_month)

```


CAMHS
=======================================================================

Row
-----------------------------------------------------------------------

### People in contact with CAMHS

```{r}
#Extract number
cyp01_number <- reactive({
  MHSDS_main_reac_both() %>%
  filter(.,MEASURE_ID=="CYP01") %>%
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  slice_max(end_date) %>%
  pull(MEASURE_VALUE)
})
#Show number
renderValueBox({
  cyp01_value <- formatC(cyp01_number(), digits = 0, format = "f",big.mark=",")
  valueBox(
    value = cyp01_value,
    icon = "fa-child",
    color = "primary"
  )
})
```

### New referrals to CAMHS

```{r}
#Extract number
cyp32_number <- reactive({
  MHSDS_main_reac_both() %>%
  filter(.,MEASURE_ID=="CYP32") %>%
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  slice_max(end_date) %>%
  pull(MEASURE_VALUE)
})
#Show number
renderValueBox({
  cyp32_value <- formatC(cyp32_number(), digits = 0, format = "f",big.mark=",")
  valueBox(
    value = cyp32_value,
    icon = "fa-arrow-up",
    color = "primary"
  )
})
```

### Open ward stays

```{r}
#Extract number
cyp21_number <- reactive({
  MHSDS_main_reac_both() %>%
  filter(.,MEASURE_ID=="CYP21") %>%
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  slice_max(end_date) %>%
  pull(MEASURE_VALUE)
})
#Show number
renderValueBox({
  cyp21_value <- formatC(cyp21_number(), digits = 0, format = "f",big.mark=",")
  valueBox(
    value = cyp21_value,
    icon = "fa-hospital",
    color = "primary"
  )
})
```

Row
-----------------------------------------------------------------------

### People in contact with CAMHS {data-width=500}

```{r}

output$people_in_contact <- renderPlotly({
  
CAMHS_contacts_data <- MHSDS_main_reac() %>%
  filter(MEASURE_ID %in% c("CYP01","CYP32")) %>%
  select(.,start_date,end_date,PRIMARY_LEVEL_DESCRIPTION,MEASURE_ID,MEASURE_VALUE) %>% 
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE),
         timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID"),
         type=ifelse(MEASURE_ID=="CYP01","In contact with CAMHS","New referrals"),
         MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID"))) %>% 
  arrange(.,start_date) %>%
  as_tibble()

CAMHS_contacts_chart <- CAMHS_contacts_data %>%
  ggplot(., aes(x=start_date, y=MEASURE_VALUE, group=type)) +
  geom_line(aes(color=type),size=1) +
  scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~timing, scales = "free_x") +
  theme_ipsum() +
  xlab("") +
  ylab("") +
  labs(col="") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))
  
ggplotly(CAMHS_contacts_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('people_in_contact', width = "100%")

```

### Crisis referrals (<18 years) {data-width=500}

```{r}

output$crisis_contacts <- renderPlotly({
  
urgent_crisis_referrals_data <- MHSDS_main_reac() %>%
  filter(MEASURE_ID %in% c("CCR70b","CCR71b")) %>%
  mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID"),
         type=ifelse(MEASURE_ID=="CCR70b","emergency","urgent"),
         MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID"))) %>%
  select(.,PRIMARY_LEVEL_DESCRIPTION,start_date,end_date,month_year,timing,type,MEASURE_VALUE)

urgent_crisis_referrals_chart <- urgent_crisis_referrals_data %>%
  ggplot(.) +
  geom_bar(aes(x=start_date, y=MEASURE_VALUE, fill=type), position="stack", stat="identity") +
  scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~timing, scales = "free_x") +
  theme_ipsum() +
  xlab("") +
  ylab("") +
  labs(fill="") +
  scale_fill_manual(values=c("urgent" = "brown", "emergency" = "firebrick2")) +
  theme(panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

ggplotly(urgent_crisis_referrals_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))
  
})

plotlyOutput('crisis_contacts', width = "100%")

```

Consultation mode
=======================================================================

Row
-----------------------------------------------------------------------

### Number of mental health contacts by mode {data-width=500}

```{r}

output$contacts_by_mode <- renderPlotly({

consultation_mode_data <- MHSDS_main_reac()  %>%
  filter(.,MEASURE_ID=="MHS30e") %>%
  filter(.,!(SECONDARY_LEVEL_DESCRIPTION %in% c("Invalid","Missing"))) %>%
  mutate(.,mode_cat=fct_collapse(SECONDARY_LEVEL_DESCRIPTION,
                                 F2F=c("Face to face communication"),
                                 Virtual=c("Telephone","Telemedicine web camera"),
                                 `E-mail/SMS`=c("Email","Short Message Service (SMS) - Text Messaging"),
                                 Other=c("Talk type for a person unable to speak","Other"))) %>%
  mutate(mode_cat=fct_relevel(mode_cat, c("E-mail/SMS","Other","Virtual","F2F"))) %>% 
  group_by(PRIMARY_LEVEL_DESCRIPTION,start_date,end_date,month_year,mode_cat) %>%
  summarise(MEASURE_VALUE=sum(as.numeric(MEASURE_VALUE),na.rm=TRUE)) %>% 
  ungroup() %>%
  mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID")) %>%
  mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID")))

consultation_mode_chart <- consultation_mode_data %>%
  ggplot(.) +
  geom_bar(aes(x=start_date, y=MEASURE_VALUE, fill=mode_cat), position="stack", stat="identity") +
  scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~timing, scales = "free_x") +
  theme_ipsum() +
  xlab("") +
  ylab("") +
  labs(fill="Mode") +
  scale_fill_brewer(palette = "Set1") +
  theme(panel.border = element_blank(),
        text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 12),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 12),
        axis.title.y = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

ggplotly(consultation_mode_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('contacts_by_mode', width = "100%")

```

Eating disorders
=======================================================================

Row
-----------------------------------------------------------------------

### Number of people waiting for treatment (<18 yrs) {data-width=500}

```{r}

output$ed_waiting <- renderPlotly({

number_waiting_ed_data <- MHSDS_ED_reac() %>%
  filter(.,MEASURE_ID %in% c("ED88","ED89")) %>%
  select(.,start_date,end_date,PRIMARY_LEVEL_DESCRIPTION,MEASURE_ID,MEASURE_VALUE) %>%
  pivot_wider(names_from = MEASURE_ID,
              names_sep = "_",
              values_from = MEASURE_VALUE) %>%
  rename(.,urgent="ED89",
         all="ED88") %>%
  mutate(.,`non urgent`=as.numeric(all)-as.numeric(urgent),
         all=as.numeric(all),
         urgent=as.numeric(urgent)) %>%
  pivot_longer(
    cols = all:`non urgent`,
    names_to = c("type"),
    values_to = "count"
  ) %>%
  filter(.,type!="all") %>%
  mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID")) %>%
  mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID")))

number_waiting_ed_chart <- number_waiting_ed_data %>%
  ggplot(.) +
  geom_bar(aes(x=start_date, y=count, fill=type), position="stack", stat="identity") +
  scale_x_date(date_labels = "%b %Y",date_breaks = "1 month") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~timing, scales = "free_x") +
  theme_ipsum() +
  xlab("") +
  ylab("") +
  labs(fill="") +
  scale_fill_manual(values=
                      c("urgent" = "brown", "non urgent" = "darkseagreen4")) +
  theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))

ggplotly(number_waiting_ed_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('ed_waiting', width = "100%")

```

### Number of people starting treatment within target time {data-width=500}

```{r}

output$ed_target_time <- renderPlotly({

target_time_ed_data <- MHSDS_ED_reac() %>%
  filter(.,MEASURE_ID %in% c("ED86e","ED87e")) %>%
  select(.,start_date,end_date,PRIMARY_LEVEL_DESCRIPTION,MEASURE_ID,MEASURE_VALUE) %>%
  mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID")) %>%
  mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID"))) %>%
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE),
         Type=case_when(MEASURE_ID=="ED86e" ~ "urgent",
                        MEASURE_ID=="ED87e" ~ "non urgent",
                        TRUE ~ "NA"))

target_time_ed_chart <- target_time_ed_data %>%
  ggplot(., aes(x=start_date, y=MEASURE_VALUE, group=Type)) +
  geom_line(aes(color=Type),size=1) +
  scale_x_date(date_labels = "%b %Y",date_breaks = "1 month") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~timing, scales = "free_x") +
  theme_ipsum() +
  xlab("") +
  ylab("") +
  labs(col="") +
  scale_colour_manual(values=
                      c("urgent" = "brown", "non urgent" = "darkseagreen4")) +
  theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))

ggplotly(target_time_ed_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('ed_target_time', width = "100%")
#tableOutput("ed_target_time")

```

Other metrics
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### Number of discharges from referral (<18 yrs) {data-width=500}

```{r}

output$discharges <- renderPlotly({

discharge_data <- MHSDS_main_reac()  %>%
   filter(.,MEASURE_ID %in% c("MHS57a")) %>%
   mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID"),
          MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
   mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID"))) %>%
   select(.,PRIMARY_LEVEL_DESCRIPTION,start_date,end_date,month_year,timing,MEASURE_VALUE)

discharge_chart <- discharge_data %>%
   ggplot(.) +
   geom_bar(aes(x=start_date, y=MEASURE_VALUE), position="stack", stat="identity",fill="cornflowerblue") +
   scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
   scale_y_continuous(labels = scales::comma) +
   facet_wrap(~timing, scales = "free_x") +
   theme_ipsum() +
   xlab("") +
   ylab("") +
   theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))

ggplotly(discharge_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('discharges', width = "100%")

```

### Number of ward stays >50km away {data-width=500}

```{r}

output$far_ward <- renderPlotly({

far_ward_data <- MHSDS_main_reac() %>%
   filter(.,MEASURE_ID %in% c("MHS22a")) %>%
   mutate(.,timing=ifelse(start_date<ymd("2020-04-01"),"Pre-COVID","Post-COVID"),
          MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
   mutate(.,timing=fct_relevel(timing, c("Pre-COVID","Post-COVID"))) %>%
   select(.,PRIMARY_LEVEL_DESCRIPTION,start_date,end_date,month_year,timing,MEASURE_VALUE)
 
far_ward_chart <- far_ward_data %>%
   ggplot(.) +
   geom_bar(aes(x=start_date, y=MEASURE_VALUE), position="stack", stat="identity",fill="darkslateblue") +
   scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
   scale_y_continuous(labels = scales::comma) +
   facet_wrap(~timing, scales = "free_x") +
   theme_ipsum() +
   xlab("") +
   ylab("") +
   theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))
 
 ggplotly(far_ward_chart) %>%
  layout(autosize = T, margin = m) %>%
  layout(legend = list(orientation = 'h'))

})

plotlyOutput('far_ward', width = "100%")

```

Row {data-height=600}
-----------------------------------------------------------------------

### Changes in activity compared to previous year {data-width=500}

```{r}

output$pct_change <- renderPlotly({

CAMHS_people_in_contact_data <- MHSDS_main_reac() %>%
  filter(.,MEASURE_ID %in% c("CYP01","CYP23","MHS61a")) %>%
  select(.,PRIMARY_LEVEL_DESCRIPTION,MEASURE_ID,MEASURE_VALUE,start_date,month_year) %>%
  mutate(.,MEASURE_VALUE=as.numeric(MEASURE_VALUE),
         start_date_l1=start_date-years(1),
         MEASURE_KEY=case_when(
           MEASURE_ID=="CYP01" ~ "People in contact",
           MEASURE_ID=="CYP23" ~ "Open referrals",
           MEASURE_ID=="MHS61a" ~ "First contacts (<18)",
           TRUE ~ "NA"
         ))

CAMHS_people_in_contact_l1 <- CAMHS_people_in_contact_data %>%
  select(.,start_date,MEASURE_VALUE,PRIMARY_LEVEL_DESCRIPTION,MEASURE_ID,MEASURE_KEY) %>%
  rename(.,MEASURE_VALUE_l1=MEASURE_VALUE,start_date_l1=start_date)

CAMHS_people_in_contact_data <- left_join(CAMHS_people_in_contact_data,
          CAMHS_people_in_contact_l1,
          by=c("start_date_l1","PRIMARY_LEVEL_DESCRIPTION","MEASURE_ID","MEASURE_KEY")) %>%
  arrange(.,MEASURE_ID,start_date) %>%
  mutate(pct_change_l1=(MEASURE_VALUE-MEASURE_VALUE_l1)/MEASURE_VALUE_l1*100) %>%
  filter(.,!is.na(pct_change_l1))
rm(CAMHS_people_in_contact_l1)

CAMHS_people_in_contact_chart <- CAMHS_people_in_contact_data %>%
  ggplot(., aes(x=start_date, y=pct_change_l1, group= MEASURE_KEY)) +
  geom_line(aes(color= MEASURE_KEY),size=1) +
  scale_x_date(date_labels = "%b %Y",date_breaks = "1 month") +
  theme_ipsum() +
  xlab("") +
  ylab("% change") +
  labs(col="") +
  scale_color_manual(values=c("Open referrals" = "aquamarine4",
                              "People in contact" = "tomato3",
                              "First contacts (<18)" = "olivedrab4")) +
  theme(legend.position="bottom",
        panel.border = element_blank(),
        strip.text = element_text(size=8),
        text = element_text(size = 8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 8),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 8),
        axis.title.y = element_text(size = 8))

ggplotly(CAMHS_people_in_contact_chart) %>%
  layout(autosize = T, margin = m)
})

plotlyOutput('pct_change', width = "100%")

```

CAMHS workforce and demand
=======================================================================

Row
-----------------------------------------------------------------------

### People in contacts with CAMHS compared to FTE staff {data-width=500}

```{r}

output$contactsvsstaff <- renderPlotly({

############## CAMHS activity level

CAMHS_contacts <- MHSDS_main_pooled_dashboard %>%
  filter(.,PRIMARY_LEVEL_DESCRIPTION=="England",
         MEASURE_ID=="CYP01") %>%
  select(.,start_date,MEASURE_VALUE) %>%
  mutate(.,date_ymd=lubridate::ymd(start_date)) %>%
  mutate(.,date_ymd=floor_date(date_ymd, "month"),
         measure="People in contact CAMHS",
         MEASURE_VALUE=as.numeric(MEASURE_VALUE)) %>%
  select(.,-"start_date") %>%
  arrange(.,date_ymd)

############## Consultants chart

mh_nurses_data <- NHS_workforce_other %>%
  mutate(.,date_ymd=lubridate::ymd(Date)) %>%
  mutate(.,date_ymd=floor_date(date_ymd, "month"),
         measure="FTE nurses") %>%
  rename(.,MEASURE_VALUE=`Total FTE`,
         `Specialty`=`Staff Group 2`) %>%
  filter(`Specialty` %in% c("010_Nurses - mental health")) %>%
  mutate(`Specialty`=ifelse(`Specialty`=="010_Nurses - mental health","Mental health",`Specialty`)) %>% 
  group_by(date_ymd,measure,`Specialty`) %>%
  summarise(MEASURE_VALUE=sum(MEASURE_VALUE,na.rm=TRUE)) %>% 
  ungroup()

psych_doctors_data <- NHS_workforce_doctors %>%
  mutate(.,date_ymd=lubridate::ymd(Date)) %>%
  mutate(.,date_ymd=floor_date(date_ymd, "month"),
         measure="FTE doctors") %>%
  rename(.,MEASURE_VALUE=FTE) %>%
  filter(Specialty %in% c("Child and adolescent psychiatry",
                          "General psychiatry")) %>%
  group_by(date_ymd,measure,`Specialty Group`,Specialty) %>%
  summarise(MEASURE_VALUE=sum(MEASURE_VALUE,na.rm=TRUE)) %>% 
  ungroup()

mh_all_staff_data <- mh_nurses_data %>%
  bind_rows(.,psych_doctors_data) %>% 
  bind_rows(.,CAMHS_contacts) %>% 
  mutate(.,date_ymd_l1=date_ymd-years(1))

mh_all_staff_data_l1 <- mh_all_staff_data %>%
  select(.,date_ymd,measure,`Specialty Group`,Specialty,MEASURE_VALUE) %>%
  rename(.,MEASURE_VALUE_l1=MEASURE_VALUE,date_ymd_l1=date_ymd)

mh_all_staff_data <- left_join(mh_all_staff_data,
                                mh_all_staff_data_l1,
                                by=c("date_ymd_l1","measure","Specialty Group","Specialty")) %>%
  arrange(.,Specialty,date_ymd) %>%
  mutate(pct_change_l1=(MEASURE_VALUE-MEASURE_VALUE_l1)/MEASURE_VALUE_l1*100) %>%
  mutate(.,fy_precovid=ifelse(date_ymd>=ymd("2019-04-01")&
                                date_ymd<=ymd("2020-03-01"),1,0)) %>%
  mutate(.,date_post_covid=ifelse(fy_precovid==0,date_ymd,"pre-COVID")) %>%
  mutate(.,group_for_base_chart=paste(measure,`Specialty Group`,`Specialty`,
                                      fy_precovid,date_post_covid))
rm(mh_all_staff_data_l1)


############## Level 100 chart

base100_data <- mh_all_staff_data %>%
  filter(date_ymd>=ymd("2019-04-01")) %>% 
  group_by(group_for_base_chart) %>%
  mutate(.,MEASURE_VALUE_avg=mean(MEASURE_VALUE,na.rm=TRUE)) %>%
  mutate(.,pre_value=as.numeric(ifelse(fy_precovid==1,MEASURE_VALUE_avg,NA))) %>%
  ungroup() %>%
  group_by(measure,Specialty) %>%
  mutate(.,pre_value=LOCF(pre_value)) %>%
  mutate(value_base100=MEASURE_VALUE_avg/pre_value*100) %>% 
  ungroup() %>%
  filter(date_ymd>=ymd("2020-03-01"))

base100_chart <- base100_data %>%
  mutate(chart_group=paste(measure,Specialty,sep=" ")) %>%
  mutate(chart_group=str_replace_all(chart_group," NA","")) %>% 
  ggplot(., aes(x=date_ymd, y=value_base100, group=chart_group)) +
  geom_line(aes(color=chart_group),size=1.5) +
  scale_x_date(date_labels = "%b %Y",date_breaks = "3 months") +
  scale_y_continuous(labels = scales::comma) +
  theme_ipsum() +
  xlab("") +
  ylab("Activity") +
  labs(col="",
       title = "Mental health services for children and young people", 
       subtitle = "Comparison to pre-COVID levels", 
       caption = "100 = average activity in year before March 2020") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position="right",
        panel.border = element_blank(),
        strip.text = element_text(size=12),
        text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        axis.text = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1,size = 12),
        axis.title.x = element_text(margin = unit(c(3, 0, 0, 0), "mm"),size = 12),
        axis.title.y = element_text(size = 12))

ggplotly(base100_chart) %>%
  layout(autosize = T,
         annotations = 
           list(x = 1, y = -0.1, text = "100 = average level in year pre-COVID", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=0, yshift=0,
                font=list(size=15, color="black")))
  
})

plotlyOutput('contactsvsstaff', width = "100%")

```
